### 2022-10-12 manuscript/manuscript.tex
more theory

\subsection*{Overview of Quantitative Genetic Linked Selection Theory}

Here, we step through a quick derivation of \textcite{Santiago1998-bs}.
Throughout, we assume random mating, hermaphroditic individuals, and a constant
population size. The change in a neutral allele's frequency in one generation
can be partitioned into the three sources of stochasticity: the random
associations with fitness backgrounds (i.e. \emph{draft}), the non-heritable
randomness in family size, and the Mendelian noise from heterozygotes
segregating. If we let $x_{0,i} \in \{0, \nicefrac{1}{2}, 1\}$ be the frequency
of neutral alleles individual $i$ in generation 0 carries, we can partition the
random neutral allele frequency of the population ($x_t$ without the individual
index) into the underlying stochastic causes,

@@

\begin{align}
  \label{eq:var_wf}
  \V(x_2) &= \E\left[(x_2 - x_0)^2\right] \\
          &= \E_1\left[\E_2\left[((x_2-x_1) + (x_1 - x_0))^2 | x_1\right]\right] \\
          &= \E_1\left[\E_2\left[(\Delta x_2 + \Delta x_1)^2 | x_1\right]\right] \\
          &= \E_1\left[\E_2\left[{\Delta x_2}^2|x_1\right]\right] + 2\E_1\left[\E_2[\Delta x_2 \Delta x_1 | x_1]\right] + \E_1\left[ {\Delta x_1}^2\right] \\
          &= \E_1\left[\E_2\left[{\Delta x_2}^2|x_1\right]\right] + \E_1\left[ {\Delta x_1}^2\right] + C_{1,2} \\
          &= \frac{p_0(1-p_0)}{2N}\left(1 - \frac{1}{2N}\right) + \frac{p_0(1-p_0)}{2N} + C_{1,2}
\end{align}

Consequently, the variance of the neutral allele frequency changes each
generation according to the probability of failing to coalesce each generation
and the pairwise covariance terms $C_{i,j}$ that build up due to associations
with heritable fitness backgrounds. Note that the covariance term $C_{1,2} =
\E_1\left[\E_2[\Delta x_2 \Delta x_1 | x_1]\right] = 0$ under neutral
evolution, and this simplifies to the well-known equation for variance in a
Wright--Fisher population.

Now, we turn our attention to the case where there is heritable fitness
variation, which leads to non-zero covariance terms $C_{i,j}$. These terms
emerge when $\var(k)$ has a heritable component of fitness that can be
transmitted along with the neutral allele, thereby affecting the neutral
allele's trajectory in later generations. If we write the offspring number into
heritable and non-heritable components, $k_i = 2f_i + \varepsilon_i$, then
$\var(k_i) = 4 V_h + V_n$. We can further decompose the $K_1$ term into $K_1 =
S_1 + D_1$,

@@

In general, the evolution of the variance in allele frequency change in a
system with polygenic selection is given by,

@@

Since the probability of coalesce (or, equivalently, identity by descent) is
proportional to the variance in allele frequency change
\parencite{Barton2000-zg}, this encodes the pairwise coalescent rate of the
population through time under polygenic selection through the $N_{e,t}$ terms.
In the original derivation of \textcite{Santiago1995-hx}, they set each
$N_{e,t}$ term to a single $N$ and solve this summation in the limit to
infinity giving,

@@

(cf. equation 18, \cite{Santiago1995-hx}). Levels of pairwise diversity,
however, depend on the full distribution of pairwise coalesce times, or the
$N_{e,t}$. While this would require solving the recursion defined by equations
\eqref{eqn:var_xt} and \eqref{eqn:ne_var}, \textcite{Santiago1998-bs} show
that a good approximation is just assume a single effective population size for
all timepoints, $N_{e,t} = N_e$, but use the time-indexed $Q_t$. This leads to,

\begin{align}
  N_{e,t} &\approx \frac{4N}{4 V_h Q_t^2 +V_n+2}
\end{align}

% HERE

@@

There has been considerable past work to find both the rate of the ratchet
(XXX), and summaries of the genealogies created under weak purifying selection
(XXX). Our approximation combines a few theoretic avenues from past work. Our
primary goal is to find a theoretic scaling function $B$ that rescales the neutral
pairwise coalescent times the appropriate amount to

Strictly speaking, there is no way to rescale the coalescent under general
background selection process, as genealogies deviate from the Kingman
coalescent slightly under strong selection
\parencite{Walczak2012-fi,Cvijovic2018-vd} and greatly under weak selection
\parencite{OFallon2010-my}

First, we approximate the coalescent effective population size, which
determines expected pairwise diversity, from the variance effective population
size. Then, we use a general quantitative genetic model of linked selection to
find the variance effective population size under background selection. Since
the impact of selection under this model is parameterized by the additive
genetic fitness variance, we use an approximation for this based on the
additive \emph{genic} variance, which depends on the rate of the ratchet.
Previous work has been unable to find a simple closed-form solution to the rate
of the ratchet because it depends on moments of the fitness distribution (e.g.
fitness variation), which in turn depend on even higher-moments of the fitness
distribution (e.g. the skew of the fitness distribution). While the ratchet
should scale with the probability of fixation, the probability of fixation
(Kimura, Malecot, XXX) is not accurate when multiple deleterious mutations are
segregating simultaneously and in close linkage with one another. A key insight
of \textcite{Santiago2016-mu} was that the classic equation for the probability
of fixation is fairly robust as long as the correct effective population size
is used. Their approach is to solve a system of two non-linear equations: one
for the variance effective population size (which depends on the rate of the
ratchet), and another for the time to fixation (which depends on the effective
population size).

Below, we step through the theory we use in our statistical methods. We
re-derive central parts to illustrate where approximations have been made,
which allows us to understand how these approximations may impact our
statistical methods and inspect model fit in regions where these approximations
are less likely to hold.

\subsection*{Pairwise Diversity under Background Selection}

The central quantity of interest for our statistical inference methods is
average pairwise diversity along the chromosome. While other summaries of the
data, such as the allele frequency spectrum, may be more informative about the
evolutionary processes acting in a region, we lack population genetic theory
for these quantities under most selective regimes and they require more degrees
of freedom to characterize. By contrast, neutral diversity under an
infinite-sites model is determined only by the per-basepair mutation rate $\mu$
and the expected pairwise coalescent time $\E[T_2]$,

\begin{align}
  \E(\pi) = 2 \mu \E(T_2)
\end{align}

Let us assume that there is a constant \emph{drift-effective} population size
$N_d$ which determines the rate of coalescence due only to neutral demographic
processes. Additionally, we assume some background selection process that in
expectation scales the effective population size by $b(t)$ at time $t$ (though
this could in theory be due to any process). Then, the expected pairwise
coalescent time is

\begin{align}
  \label{eq:T_2}
  \E(T_2) = \sum_{i=1}^\infty \frac{i}{2 b(i) N_d}\prod_{t=1}^{i-1} \left(1-\frac{1}{2 b(t) N_d}\right).
\end{align}

The time-dependency of $b(t)$ is required because the background selection
process will usually alter the rate of coalesce through time. For example,
under weak and intermediate selection, coalesce rates speed up as lineages move
in the past under intermediate to weak selection, similar to population growth
\parencite{Walczak2012-fi,Seger2010-rs,Good2014-yz}.

% Intuitively, this is
% because lineages are unlikely to coalesce with numerous less fit lineages they
% encounter as they go back in time, until ultimately coalescing in one of the
% least-loaded fitness classes.

This formulation is general and classic background selection theory (in the
no-ratchet domain) is a special case of it. Under classic BGS and load theory,
two randomly sampled lineages will each carry a number of deleterious mutations
sampled from a Poisson stationary distribution. Because any lineage carrying $j
> 0$ strongly deleterious mutations is destined to extinction, these lineages
can only coalesce in a common ancestor in the least-loaded class with zero
deleterious mutations. Under the Poisson stationary distribution, the size of
this class is $f_0 N$, where $f_0 = e^{-\nicefrac{U}{s}}$, so the coalescent
rate is scaled by $\nicefrac{1}{f_0}$. Once the lineages reach this fitness
class, there are no fitness differences among individuals, so they coalesce
following the Kingman process. A lineage moves backwards in time $t_d$
generations through the ``delay phase" to the least-loaded class. Classic BGS
theory ignores this delay phase, since $t_d \sim \log(j+1)/s << f_0 N$
(Durrett, p. 213).

%Here, we model the background selection process forward in time. We approximate
%the expected pairwise diversity with the probability of heterozygosity,
%$\E(\pi) \approx \mathcal{H}$, which holds under weak mutation and random
%mating, such that

%\begin{align}
%  %\mathcal{H} =  2 \mu \sum_{i=1}^\infty \prod_{t=1}^i \left(1-\frac{1}{2 b(t) N_d} \right)
%  \mathcal{H} &= 1 - \sum_{i=1}^\infty (1-\mu)^{2i} \prod_{t=1}^i \left(1-\frac{1}{2 b(t) N_d} \right) \\
%              &\approx 2 \mu \sum_{i=1}^\infty \prod_{t=1}^i \left(1-\frac{1}{2 b(t) N_d} \right)
%\end{align}

%in the limit where $\mu \to 0$. This is equation 4 of
%\textcite{Santiago2016-mu}; they define the sum as the heterozygosity effective
%population size.

\subsection*{Forward in Time Models of Linked Selection}

Next, show how quantitative genetic models of linked selection
\parencite{Robertson1961-ho,Santiago1995-hx,Santiago1998-bs,Santiago2016-mu}
find expressions for $b(t)$ with by modeling the selective process forward in
time. Their approach is to consider how the variance in neutral allele
frequency change is impacted by a linked selection process.

\subsubsection*{Variance in Neutral Allele Frequency in a Neutral System}

For comparison, we start with the recursion for the variance in allele
frequency change in a neutral Wright--Fisher model. Under binomial sampling of
alleles, the random frequency $x_t$ is $2N x_t \sim \mathrm{Binom}(x_{t-1},
2N)$. If we look forward in time on generation and treat $x_0$ as fixed,

\begin{align}
  \V(x_2) &= \E\left[(x_2 - x_0)^2\right] \\
          &= \E_1\left[\E_2\left[((x_2-x_1) + (x_1 - x_0))^2 | x_1\right]\right] \\
          &= \E_1\left[\E_2\left[(\Delta x_2 + \Delta x_1)^2 | x_1\right]\right] \\
          &= \E_1\left[\E_2\left[{\Delta x_2}^2|x_1\right]\right] + 2\E_1\left[\E_2[\Delta x_2 \Delta x_1 | x_1]\right] + \E_1\left[ {\Delta x_1}^2\right] \\
          &= \E_1\left[\E_2\left[{\Delta x_2}^2|x_1\right]\right] + \E_1\left[ {\Delta x_1}^2\right] \\
          &= \frac{x_0(1-x_0)}{2N}\left(1 - \frac{1}{2N}\right) + \frac{x_0(1-x_0)}{2N}
\end{align}

Note that the cross-moment $\E_1\left[\E_2[\Delta x_2 \Delta x_1 | x_1]\right]
= 0$ under neutral evolution. In general, the variance in allele frequency
after $t$ generations of neutral evolution is,

\begin{align}
  \label{eq:var_freq}
  %\V(x_t) &= \frac{x_0(1-x_0)}{2N}\left(\sum_{i=1}^t \left(1 - \frac{1}{2N}\right)^{i-1} \right)
  \V(x_t) &= \V(x_1) \left(1 + \sum_{i=1}^t \left(1 - \frac{1}{2N}\right)^{i} \right)
\end{align}

This geometric series converges to the familiar expression,

\begin{align}
  \V(x_t) &= x_0(1-x_0)\left(1 - \left(1 - \frac{1}{2N}\right)^t \right).
\end{align}

which in the limit $t \to \infty$, converges to the Bernoulli variance
$x_0(1-x_0)$ as the frequency either goes to 1 with probability $x_0$ or 0 with
probability $1-x_0$.

% Now, if the effective population size were to change each generation, equation
% \eqref{eq:var_freq} would be,

% \begin{align}
%   \V(x_t) &= \V(x_0) \left(1 + \sum_{i=1}^t \prod_{j=1}^{i} \left(1 - \frac{1}{2N(j)}\right) \right)
% \end{align}

% Note that this too converges to $x_0(1-x_0)$ as $t \to \infty$.

\subsubsection*{Variance in Neutral Allele Frequency with Linked Selection}

Note that equation XXX is a special case, when the evolutionary process has
independent allele frequency changes through the generations and thus the
cross-moment is zero. Selection generates correlations between allele frequency
changes at nearby neutral sites. These correlations created by linked selection
inflate the variance in neutral allele frequency $\V(x_t)$ beyond what it would
be in a system without linked selection operating. The magnitude of this
inflation term is set by the selection process and its dynamics, linkage
disequilibrium, and the recombination rate. Even k

The theory of this began with \textcite{Robertson1961-ho},
and was thoroughly worked out by Santiago and Caballero
(\citeyear{Santiago1995-hx}, \citeyear{Santiago1998-bs}; see also
\cite{Barton2000-zg} and \cite{Buffalo2019-qs}). Our BGS statistical methods
employ the extension of this model for weak BGS \parencite{Santiago2016-mu}.
For completeness, we step through a quick derivation of the linked selection
inflation factor models here, based on
\textcite{Santiago1995-hx,Santiago1998-bs}.

Quantitative genetic models of linked selection

%%%%%

One way we can model these dependencies is to partition $\Delta_i$ into drift
and selection components. Over evolutionary replicates, the drift components
are independent and the selection components have some degree of
autocorrelation. Let us partition $\Delta x_i = \Delta_S x_i + \Delta_D x_i$,

\begin{align}
  \V(x_2) &= \V\left(\Delta x_2 + \Delta x_1\right) \\
          &= \V(\Delta_D x_2) + \V(\Delta_D x_1) + 2 \cov(\Delta_S x_2, \Delta_S x_1) \\
          &= \frac{x_0(1-x_0)}{2N}\left(1 - \frac{1}{2N}\right) + \frac{x_0(1-x_0)}{2N} +  2 \cov(\Delta_S x_2, \Delta_S x_1) 
\end{align}

since $\Delta_D x_i \perp \Delta_D x_j$ for $i \ne j$. The covariances are
determined by the selective process dynamics, linkage disequilibrium, and the
recombination rate. These covariances are throughly worked out in
\textcite{Santiago1995-hx} and \textcite{Buffalo2019-qs}, so we give a short
derivation here. Throughout, we assume random mating, hermaphroditic
individuals, and a constant population size. We will use angular brackets to
indicate an average over all diploid individuals in the population, of some
value with an individual-level index. For example, if we denote the number of
allele an individual carries in generation $t$ as $z_{t,i} \in \{0, 1, 2\}$,
the population frequency at time $x_t$ is

\begin{align}
  x_t = \langle z_t \rangle = \frac{1}{2N} \sum_{i=1}^N z_{t,i}
\end{align}

>>>>

\subsection*{Overview of Quantitative Genetic Models of Effective Population Size}

Here, we step through a quick derivation of \textcite{Santiago1995-hx}.
Throughout, we assume random mating, hermaphroditic individuals, and a constant
population size. The change in a neutral allele's frequency in one generation
can be partitioned into the three sources of stochasticity: the random
associations with fitness backgrounds (i.e. \emph{draft}), the non-heritable
randomness in family size, and the Mendelian noise from heterozygotes
segregating. If we let $x_{0,i} \in \{0, \nicefrac{1}{2}, 1\}$ be the frequency
of neutral alleles individual $i$ in generation 0 carries, we can partition the
random neutral allele frequency of the population ($x_t$ without the individual
index) into the underlying stochastic causes,

@@

\begin{align}
  \label{eq:var_wf}
  \V(x_2) &= \E\left[(x_2 - x_0)^2\right] \\
          &= \E_1\left[\E_2\left[((x_2-x_1) + (x_1 - x_0))^2 | x_1\right]\right] \\
          &= \E_1\left[\E_2\left[(\Delta x_2 + \Delta x_1)^2 | x_1\right]\right] \\
          &= \E_1\left[\E_2\left[{\Delta x_2}^2|x_1\right]\right] + 2\E_1\left[\E_2[\Delta x_2 \Delta x_1 | x_1]\right] + \E_1\left[ {\Delta x_1}^2\right] \\
          &= \E_1\left[\E_2\left[{\Delta x_2}^2|x_1\right]\right] + \E_1\left[ {\Delta x_1}^2\right] + \mathcal{C}_{1,2} \\
          &= \frac{p_0(1-p_0)}{2N}\left(1 - \frac{1}{2N}\right) + \frac{p_0(1-p_0)}{2N} + \mathcal{C}_{1,2}
\end{align}

Consequently, the variance of the neutral allele frequency changes each
generation according to the probability of failing to coalesce each generation
and the pairwise covariance terms $\mathcal{C}_{i,j}$ that build up due to associations
with heritable fitness backgrounds. Note that the covariance term $\mathcal{C}_{1,2} =
\E_1\left[\E_2[\Delta x_2 \Delta x_1 | x_1]\right] = 0$ under neutral
evolution, and this simplifies to the well-known equation for variance in a
Wright--Fisher population.

Now, we turn our attention to the case where there is heritable fitness
variation, which leads to non-zero covariance terms $\mathcal{C}_{i,j}$. These
terms emerge when $\var(k)$ has a heritable component of fitness that can be
transmitted along with the neutral allele, thereby affecting the neutral
allele's trajectory in later generations. If we partition the offspring number
into heritable and non-heritable components, $k_i = 2f_i + \varepsilon_i$, then
$\var(k_i) = 4 V_h + V_n$. When there is heritable fitness variation across
individuals, $V_h > 0$. Because the population size is assumed to be constant,
the mean heritable fitness across individuals is constrained to be $\E_i(f_i) =
1$; this implies that the individual fitness values $f_i$ are relative
fitnesses as is standard in population genetics, in this case to the population
mean. Thus $V_h$ is the squared coefficient of heritable fitness variation
(this is often denoted as $C^2 = \nicefrac{V_A}{\bar{w}^2}$, see
\cite{Crow1958-pc,Charlesworth1987-ab,Houle1992-ur}).

We can further decompose the $K_1$ term into $K_1 = S_1 + D_1$,

@@

In general, the evolution of the variance in allele frequency change in a
system with heritable fitness due to a single locus $r$ recombination fraction
away from the neutral site is given by,

@@

Since the probability of coalesce (or, equivalently, identity by descent) is
proportional to the variance in allele frequency change
\parencite{Barton2000-zg}, this encodes the pairwise coalescent rate of the
population through time under selection through the $N_{e,t}$ terms.  In the
original derivation of \textcite{Santiago1995-hx}, they set each $N_{e,t}$ term
to a single $N$ and solve this summation in the limit to infinity giving,

@@

(cf. equation 18, \cite{Santiago1995-hx}). Levels of pairwise diversity,
however, depend on the full distribution of pairwise coalesce times, or the
$N_{e,t}$. While this would require solving the recursion defined by equations
\eqref{eqn:var_xt} and \eqref{eqn:ne_var}, \textcite{Santiago1998-bs} show that
a reasonably good approximation follows from using a single effective
population size for all timepoints, $N_{e,t} = N_e$, but use the time-indexed
$Q_t$. This leads to,

\begin{align}
  \label{eq:ne_t}
  N_{e,t} &\approx \frac{4N}{4 V_h Q_t^2 +V_n+2}
\end{align}

(c.f. equation 15 of \cite{Santiago1995-hx} and p. 1270 of \cite{Santiago2016-mu}).

\subsection*{Effective Population Size Under Polygenic Selection}

The results of the previous section assume only a single locus a recombination
fraction $r$ apart from the neutral locus is determining the coalesce rate. The
model of \textcite{Santiago1998-bs} extends this to the case where fitness is
polygenic. As they show in their original paper, this system can accommodate a
variety of different equilibrium selection systems as long as expressions for
$V_h$ and $Z$ can be worked out. Other authors have extended similar models to
more elaborate bbreeding structures \parencite{Wray1990-zf,Woolliams1993-qo}.

First, we extend equation \eqref{eq:ne_t} to a polygenic system. Throughout, a
multiplicative fitness model is assumed (i.e. independent effects, no
epistasis), where the fitness of individual $i$ is the product of their fitness
contributions from each locus $l$, $w_{i,l}$, giving $w_i = \prod_{l=1}^L
w_{i,l}$. Then, assuming mean fitness is one, the total fitness variation is
multiplicative across the locus-level fitnesses,

\begin{align}
  V_h %&= \E_i\left( w_i^2 \right) - 1 \\
      %&= \prod_{l=1}^L (1 + \alpha_{l}^2 \V(g_l))  - 1  \\
      &= \prod_{l=1}^L (1 + V_{h,l})  - 1
\end{align}

Now, the $Q_t^2 V_h$ term in the single-locus model is now under a
multiplicative polygenic model,

\begin{align}
  Q_t^2 V_h &= \prod_{l=1}^L \left(1 + \frac{V_{h,l} Q_{t,l}^2}{2}\right)  - 1
\end{align}

where $Q_{t,l}$ is the cumulative impact of selection due to the selected site
$l$ (c.f. equation 4 \cite{Santiago1998-bs}). The factor of $\nicefrac{1}{2}$
is due to the fact that we are ignoring the chance that the fitness background
on the homologous chromosome recombines onto the haplotype with the tracked
neutral allele (i.e. due to the $Q_t''$ associations). In other words, only
half the fitness variation in a diploid can stay associated with the neutral
allele through the generations.

To simplify the derivation, assume $V_n = 2$ as under a Wright--Fisher model.
Then, the effective population size is,

\begin{align}
  N_{e,t} &\approx \frac{N}{V_h Q_t^2 + 1} \\
          &\approx \frac{N}{\prod_{l=1}^L \left(1 + \frac{V_{h,l} Q_{t,l}^2}{2}\right)} \\
          &\approx N \exp\left(- \sum_{l=1}^L \frac{V_{h,l} Q_{t,l}^2}{2}\right)
\end{align}

(c.f. \cite{Santiago1998-bs} equation 4b). This summation can then in turn be
approximated by a integral; see \textcite{Santiago1998-bs} for more. Note that
another way to arrive at this equation is to take the product of reduction
factors $\nicefrac{N_e}{N}$ across selected sites, as in equation 4 of
\textcite{Hudson1995-xc}.

The heritable variance in fitness $V_h$ associated with the tracked neutral
allele does not remain constant through the generations, but decays due to
selection and drift, and is maintained by the input of new mutations. For an
arbitrary polygenic selective system, the dynamics are extraordinarily
complicated due to the complexity of multilocus selection. However, several
quantitative genetic models simplify these dynamics by assuming an
infinitesimal model, where the trait is sufficiently polygenic that the allele
frequency changes at the underlying loci due to selection are negligible
compared to drift. This allows for a critical assumption of the quantitative
genetic models of linked selection, that the decay rate in variation is
constant (and thus independent of the underlying allele frequency changes). In
the Santiago and Caballero model, this is the decay factor $Z$.

The theory of \textcite{Santiago1998-bs} is based on an equilibrium model,
where $\Delta V_h = 0$ of the long run, even if the fitness variation
\emph{associated} with the allele's haplotype is changing. This macro-scale

\begin{align}
  \Delta V_h &= V_m - V_h\left(1 - \frac{1}{2 N_e}\right) \\
\end{align}

\subsection*{Effective Population Size Under Strong Background Selection}

Under strong background selection, multiple segregating deleterious sites
contribute to fitness variation. BGS models assume a multiplicative fitness
model, such that the fitness of an individual $i$ is

\begin{align}
  w_i = (1-s)^{n_i}
\end{align}

where $n_i$ is the number of deleterious mutations this individual carries.

Classic background selection theory assumes an infinite population size, such
that mutation frequencies are at their mutation-selection equilibrium and the
expected number of mutations is $n_i = \nicefrac{\mu}{s}L$ where $L$ is the
total number of basepairs that can be mutated (the mutational target size).
Under a multiplicative fitness function in an infinite population, selection
cannot generate linkage disequilibria \parencite{Turelli1990-kd}. Thus the
classic strong background selection model assumes additive genetic fitness is
equal to the additive genic fitness (which excludes the contribution of
covariance between selected sites),

\begin{align}
  \V_i(w_i) &= 2 \sum_{l=1}^L p(1-p) s^2 \\
            &= 2 L \frac{\mu}{s} \left( 1 - \frac{\mu}{s}\right) s^2 = U s + \mathcal{O}(\mu^2) \approx Us.
\end{align}

where we define $U := 2 L \mu$ to be the deleterious mutation rate per diploid
genome, per generation.

Next, \textcite{Santiago1998-bs} find a general
equation for the factor $Z$ by which fitness variation associated with the
neutral allele is reduced under mutation-selection balance. This is a
macro-scale variable reflecting the average balance between the average loss of
fitness variance due to selection under an infinitesimal model and drift. Each
generation, the gain of new fitness variance per generation is $V_m$.

\begin{align}
  Z = 
\end{align}

Then, with under the strong background selection model,
a site under purifying selection reduces diversity at a focal neutral site $r$
by a factor of

\begin{align}
  \frac{N_{e,t}}{N} = \exp \left(-\frac{V_{h,1} Q_{t,1}^2}{2}\right)
\end{align}

@@

%There has been considerable past work to find both the rate of the ratchet
%(XXX), and summaries of the genealogies created under weak purifying selection
%(XXX). Our approximation combines a few theoretic avenues from past work. Our
%primary goal is to find a theoretic scaling function $B$ that rescales the neutral
%pairwise coalescent times the appropriate amount to

%Strictly speaking, there is no way to rescale the coalescent under general
%background selection process, as genealogies deviate from the Kingman
%coalescent slightly under strong selection
%\parencite{Walczak2012-fi,Cvijovic2018-vd} and greatly under weak selection
%\parencite{OFallon2010-my}

%First, we approximate the coalescent effective population size, which
%determines expected pairwise diversity, from the variance effective population
%size. Then, we use a general quantitative genetic model of linked selection to
%find the variance effective population size under background selection. Since
%the impact of selection under this model is parameterized by the additive
%genetic fitness variance, we use an approximation for this based on the
%additive \emph{genic} variance, which depends on the rate of the ratchet.
%Previous work has been unable to find a simple closed-form solution to the rate
%of the ratchet because it depends on moments of the fitness distribution (e.g.
%fitness variation), which in turn depend on even higher-moments of the fitness
%distribution (e.g. the skew of the fitness distribution). While the ratchet
%should scale with the probability of fixation, the probability of fixation
%(Kimura, Malecot, XXX) is not accurate when multiple deleterious mutations are
%segregating simultaneously and in close linkage with one another. A key insight
%of \textcite{Santiago2016-mu} was that the classic equation for the probability
%of fixation is fairly robust as long as the correct effective population size
%is used. Their approach is to solve a system of two non-linear equations: one
%for the variance effective population size (which depends on the rate of the
%ratchet), and another for the time to fixation (which depends on the effective
%population size).

%Below, we step through the theory we use in our statistical methods. We
%re-derive central parts to illustrate where approximations have been made,
%which allows us to understand how these approximations may impact our
%statistical methods and inspect model fit in regions where these approximations
%are less likely to hold.

%\subsection*{Pairwise Diversity under Background Selection}

%The central quantity of interest for our statistical inference methods is
%average pairwise diversity along the chromosome. While other summaries of the
%data, such as the allele frequency spectrum, may be more informative about the
%evolutionary processes acting in a region, we lack population genetic theory
%for these quantities under most selective regimes and they require more degrees
%of freedom to characterize. By contrast, neutral diversity under an
%infinite-sites model is determined only by the per-basepair mutation rate $\mu$
%and the expected pairwise coalescent time $\E[T_2]$,

%\begin{align}
%  \E(\pi) = 2 \mu \E(T_2)
%\end{align}

%Let us assume that there is a constant \emph{drift-effective} population size
%$N_d$ which determines the rate of coalescence due only to neutral demographic
%processes. Additionally, we assume some background selection process that in
%expectation scales the effective population size by $b(t)$ at time $t$ (though
%this could in theory be due to any process). Then, the expected pairwise
%coalescent time is

%\begin{align}
%  \label{eq:T_2}
%  \E(T_2) = \sum_{i=1}^\infty \frac{i}{2 b(i) N_d}\prod_{t=1}^{i-1} \left(1-\frac{1}{2 b(t) N_d}\right).
%\end{align}

%The time-dependency of $b(t)$ is required because the background selection
%process will usually alter the rate of coalesce through time. For example,
%under weak and intermediate selection, coalesce rates speed up as lineages move
%in the past under intermediate to weak selection, similar to population growth
%\parencite{Walczak2012-fi,Seger2010-rs,Good2014-yz}.

%% Intuitively, this is
%% because lineages are unlikely to coalesce with numerous less fit lineages they
%% encounter as they go back in time, until ultimately coalescing in one of the
%% least-loaded fitness classes.

%This formulation is general and classic background selection theory (in the
%no-ratchet domain) is a special case of it. Under classic BGS and load theory,
%two randomly sampled lineages will each carry a number of deleterious mutations
%sampled from a Poisson stationary distribution. Because any lineage carrying $j
%> 0$ strongly deleterious mutations is destined to extinction, these lineages
%can only coalesce in a common ancestor in the least-loaded class with zero
%deleterious mutations. Under the Poisson stationary distribution, the size of
%this class is $f_0 N$, where $f_0 = e^{-\nicefrac{U}{s}}$, so the coalescent
%rate is scaled by $\nicefrac{1}{f_0}$. Once the lineages reach this fitness
%class, there are no fitness differences among individuals, so they coalesce
%following the Kingman process. A lineage moves backwards in time $t_d$
%generations through the ``delay phase" to the least-loaded class. Classic BGS
%theory ignores this delay phase, since $t_d \sim \log(j+1)/s << f_0 N$
%(Durrett, p. 213).

%%Here, we model the background selection process forward in time. We approximate
%%the expected pairwise diversity with the probability of heterozygosity,
%%$\E(\pi) \approx \mathcal{H}$, which holds under weak mutation and random
%%mating, such that

%%\begin{align}
%%  %\mathcal{H} =  2 \mu \sum_{i=1}^\infty \prod_{t=1}^i \left(1-\frac{1}{2 b(t) N_d} \right)
%%  \mathcal{H} &= 1 - \sum_{i=1}^\infty (1-\mu)^{2i} \prod_{t=1}^i \left(1-\frac{1}{2 b(t) N_d} \right) \\
%%              &\approx 2 \mu \sum_{i=1}^\infty \prod_{t=1}^i \left(1-\frac{1}{2 b(t) N_d} \right)
%%\end{align}

%%in the limit where $\mu \to 0$. This is equation 4 of
%%\textcite{Santiago2016-mu}; they define the sum as the heterozygosity effective
%%population size.

%\subsection*{Forward in Time Models of Linked Selection}

%Next, show how quantitative genetic models of linked selection
%\parencite{Robertson1961-ho,Santiago1995-hx,Santiago1998-bs,Santiago2016-mu}
%find expressions for $b(t)$ with by modeling the selective process forward in
%time. Their approach is to consider how the variance in neutral allele
%frequency change is impacted by a linked selection process.

%\subsubsection*{Variance in Neutral Allele Frequency in a Neutral System}

%For comparison, we start with the recursion for the variance in allele
%frequency change in a neutral Wright--Fisher model. Under binomial sampling of
%alleles, the random frequency $x_t$ is $2N x_t \sim \mathrm{Binom}(x_{t-1},
%2N)$. If we look forward in time on generation and treat $x_0$ as fixed,

%\begin{align}
%  \V(x_1) = \E_1\left[(x_1 - x_0)^2\right]= \frac{x_0(1-x_0)}{2N},
%\end{align}

%and

%\begin{align}
%  \V(x_2) &= \E\left[(x_2 - x_0)^2\right] \\
%          &= \E_1\left[\E_2\left[((x_2-x_1) + (x_1 - x_0))^2 | x_1\right]\right] \\
%          &= \E_1\left[\E_2\left[(\Delta x_2 + \Delta x_1)^2 | x_1\right]\right] \\
%          &= \E_1\left[\E_2\left[{\Delta x_2}^2|x_1\right]\right] + 2\E_1\left[\E_2[\Delta x_2 \Delta x_1 | x_1]\right] + \E_1\left[ {\Delta x_1}^2\right] \\
%          &= \E_1\left[\E_2\left[{\Delta x_2}^2|x_1\right]\right] + \E_1\left[ {\Delta x_1}^2\right] \\
%          &= \frac{x_0(1-x_0)}{2N}\left(1 - \frac{1}{2N}\right) + \frac{x_0(1-x_0)}{2N}
%\end{align}

%Note that the cross-moment $\E_1\left[\E_2[\Delta x_2 \Delta x_1 | x_1]\right]
%= 0$ under neutral evolution. In general, the variance in allele frequency
%after $t$ generations of neutral evolution is,

%This geometric series converges to the familiar expression,

%\begin{align}
%  \V(x_t) &= x_0(1-x_0)\left(1 - \left(1 - \frac{1}{2N}\right)^t \right).
%\end{align}

%which in the limit $t \to \infty$, converges to the Bernoulli variance
%$x_0(1-x_0)$ as the frequency either goes to 1 with probability $x_0$ or 0 with
%probability $1-x_0$.

%% Now, if the effective population size were to change each generation, equation
%% \eqref{eq:var_freq} would be,

%% \begin{align}
%%   \V(x_t) &= \V(x_0) \left(1 + \sum_{i=1}^t \prod_{j=1}^{i} \left(1 - \frac{1}{2N(j)}\right) \right)
%% \end{align}

%% Note that this too converges to $x_0(1-x_0)$ as $t \to \infty$.

%\subsubsection*{Variance in Neutral Allele Frequency with Linked Selection}

%Note that equation XXX is a special case, when the evolutionary process has
%independent allele frequency changes through the generations and thus the
%cross-moment is zero. Selection generates correlations between allele frequency
%changes at nearby neutral sites. These correlations created by linked selection
%inflate the variance in neutral allele frequency $\V(x_t)$ beyond what it would
%be in a system without linked selection operating. The magnitude of this
%inflation term is set by the selection process and its dynamics, linkage
%disequilibrium, and the recombination rate. Even k

%The theory of this began with \textcite{Robertson1961-ho},
%and was thoroughly worked out by Santiago and Caballero
%(\citeyear{Santiago1995-hx}, \citeyear{Santiago1998-bs}; see also
%\cite{Barton2000-zg} and \cite{Buffalo2019-qs}). Our BGS statistical methods
%employ the extension of this model for weak BGS \parencite{Santiago2016-mu}.
%For completeness, we step through a quick derivation of the linked selection
%inflation factor models here, based on
%\textcite{Santiago1995-hx,Santiago1998-bs}.

%Quantitative genetic models of linked selection

%%%%%%

%One way we can model these dependencies is to partition $\Delta_i$ into drift
%and selection components. Over evolutionary replicates, the drift components
%are independent and the selection components have some degree of
%autocorrelation. Let us partition $\Delta x_i = \Delta_S x_i + \Delta_D x_i$,

%\begin{align}
%  \V(x_2) &= \V\left(\Delta x_2 + \Delta x_1\right) \\
%          &= \V(\Delta_D x_2) + \V(\Delta_D x_1) + 2 \cov(\Delta_S x_2, \Delta_S x_1) \\
%          &= \frac{x_0(1-x_0)}{2N}\left(1 - \frac{1}{2N}\right) + \frac{x_0(1-x_0)}{2N} +  2 \cov(\Delta_S x_2, \Delta_S x_1) 
%\end{align}

%since $\Delta_D x_i \perp \Delta_D x_j$ for $i \ne j$. The covariances are
%determined by the selective process dynamics, linkage disequilibrium, and the
%recombination rate. These covariances are throughly worked out in
%\textcite{Santiago1995-hx} and \textcite{Buffalo2019-qs}, so we give a short
%derivation here. Throughout, we assume random mating, hermaphroditic
%individuals, and a constant population size. We will use angular brackets to
%indicate an average over all diploid individuals in the population, of some
%value with an individual-level index. For example, if we denote the number of
%allele an individual carries in generation $t$ as $z_{t,i} \in \{0, 1, 2\}$,
%the population frequency at time $x_t$ is

%\begin{align}
%  x_t = \langle z_t \rangle = \frac{1}{2N} \sum_{i=1}^N z_{t,i}
%\end{align}
